---
title: 'STAT 1261/2260: Principles of Data Science'
subtitle: 'Homework 2'
output:
  html_document
  #word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.height=3, fig.height=5)

```


### <span style="color:green"> The Stroke Data </span>
According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

This dataset is used to predict whether a patient is likely to get stroke based on the attributes like gender, age, various diseases, and smoking status. Each row in the data provides relevant information about the patient.

#### Variables Information
##### Features (predictors):
1) `id`: unique identifier
2) `gender`: "Male", "Female" or "Other"
3) `age`: age of the patient
4) `hypertension`: 0 if the patient doesn't have hypertension, 1 if the patient has hypertension
5) `heart_disease`: 0 if the patient doesn't have any heart diseases, 1 if the patient has a heart disease
6) `ever_married`: "No" or "Yes"
7) `work_type`: "children", "Govt_jov", "Never_worked", "Private" or "Self-employed"
8) `Residence_type`: "Rural" or "Urban"
9) `avg_glucose_level`: average glucose level in blood
10) `bmi`: body mass index
11) `smoking_status`: "formerly smoked", "never smoked", "smokes" or "Unknown". "Unknown" means that the information is unavailable for this patient

##### Target (response):
12) `stroke`: 1 if the patient had a stroke or 0 if not

```{r,message=FALSE}
library(tidyverse)
data <- read_csv('strokedata.csv')
```

#### Understanding the data

```{r}
glimpse(data)
```

Note that `bmi` is coded as a character variable. Let us convert it into a numeric variable.

```{r}
data$bmi <- as.numeric(data$bmi)
```

Don't panic. These NA values were converted from "NA", the original format of missing data.

Let us check missing values for all columns.

```{r}
colSums(is.na(data))
```

Only `bmi` has missing values. Let us delete these observations for now. Later on, we will learn how to impute missing values.

```{r}
data <- data[!is.na(data$bmi),]
colSums(is.na(data))
```


```{r}
dim(data)
```

Observing the output of `glimpse()`, we see that `stroke`, `hypertension`, and `heart_disease` were coded as numeric variables. This is not ideal because they are actually binary categorical variables. Let us convert them into factor variables.

```{r}
data <- transform(data, stroke=as.factor(stroke),
                  hypertension=as.factor(hypertension),
                  heart_disease=as.factor(heart_disease))
glimpse(data[,c(4,5,12)]) # Show info about these three columns only
```

### Question 1 Categorical Variables (25 points)
General requirement: Use `ggplot2` functions only

#### 1.1 Create a horizontal bar graph for `stroke` (5 points)
Requirements:
1. Should be a horizontal bar graph
2. Add a title "Bar Graph for stroke"

```{r}
ggplot(data, aes(x = stroke)) +
  geom_bar() +
  coord_flip() +
  ggtitle("Bar Graph for stroke")
```

#### 1.2 Create a pie graph (using `ggplot2`) for gender (5 points)
Requirements:
1. Remove x and y axis labels
2. Add a title "Pie Chart for gender"

```{r}
ggplot(data, aes(x = "", fill = gender)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y") +
  labs(x = NULL, y = NULL, title = "Pie Chart for gender")
```

#### 1.3 Create a Coxcomb chart for `work_type` (5 points)

```{r}
ggplot(data, aes(x = work_type, fill = work_type)) +
  geom_bar() +
  coord_polar() +
  ggtitle("Coxcomb Chart for work_type")
```

#### 1.4 Create a bar graph for `smoking_status` and `stroke` (5 points)
Requirements:

1. Use the 'dodge' position.
2. Compare the distribution of `stroke` among different groups of `smoking_status`.
3. The x-axis label should be "Smoking Status".

```{r}
ggplot(data, aes(x = smoking_status, fill = stroke)) +
  geom_bar(position = "dodge") +
  labs(x = "Smoking Status", title = "Stroke Counts by Smoking Status")
```

#### 1.5 Creat a bar graph for `work_type` and `stroke` (5 points)
Requirements:
1. It should be a stacked bar graph that compares the distribution of work_type for different stroke categories.
2. The height of each bar should be 1.

```{r}
ggplot(data, aes(x = stroke, fill = work_type)) +
  geom_bar(position = "fill") +
  labs(y = "Proportion", title = "Distribution of work_type by stroke")
```

### Question 2 Numeric Variables (25 points)
#### 2.1 Create a histogram for `age`. (5 points)
Requirements:

1. Let the bar color be 'steelblue'
2. Let the boundary color of the bar be 'white'
3. Use 20 bins

```{r}
ggplot(data, aes(x = age)) +
  geom_histogram(bins = 20, fill = "steelblue", color = "white") +
  ggtitle("Histogram for age")
```

#### 2.2 Create a density plot to compare the distribtion of `age` for different `stroke` groups. (5 points)
Requirements:

1. Let `alpha=0.5`.
2. Let `bw=5`.

```{r}
ggplot(data, aes(x = age, fill = stroke)) +
  geom_density(alpha = 0.5, bw = 5) +
  ggtitle("Density of age by stroke")
```

#### 2.3 Create a side-by-side boxplot for `stroke` and `avg_glucose_level` (5 points)
Requirements:

1. Should be a horizontal boxplot
2. Name the x-axis(after flipping the axes) label "Average Glucose Level"

```{r}
ggplot(data, aes(x = stroke, y = avg_glucose_level)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "Average Glucose Level", title = "Average Glucose Level by stroke")
```

#### 2.4 Create a scatterplot for `bmi` and `avg_glucose_level` (5 points)
Requirements:

1. `bmi` should be the x variable.
2. Let `alpha=0.2`.
3. Use 'purple` for the points.
4. Label the y-axis "Average Glucose Level"

```{r}
ggplot(data, aes(x = bmi, y = avg_glucose_level)) +
  geom_point(alpha = 0.2, color = "purple") +
  labs(y = "Average Glucose Level", title = "bmi vs. Average Glucose Level")
```

#### 2.5 Create a scatterplot for `age` and `bmi` (5 points)
Requirements:

1. `age` should be the x variable.
2. Let `alpha=0.2`.
3. Fit and overlay a linear regression line for each "stroke" group.

```{r}
ggplot(data, aes(x = age, y = bmi, color = stroke)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("age vs. bmi by stroke group")
```
