---
title: "R Homework"
output: html_document
---

## Question 1: Object Values and Lengths

#### 1) Show the value of each object and use R command to measure the length of them.

```{r}
obj1 <- seq(20,2,-2)
obj2 <- c(2,3,5,7,11)
obj3 <- obj2 >= 5
obj4 <- list("Pizza","Pie","Cake")

obj1
length(obj1)

obj2
length(obj2)

obj3
length(obj3)

obj4
length(obj4)
```

#### 2) What values are returned by the following commands?

```{r, error=TRUE}
obj1^(-1)
obj1[-1]
obj2[4:2]
obj1 + obj2
obj1 * obj3
obj2 + obj3
obj1 + obj4
sum(obj2)
sum(obj3)
```

Notes:

- `obj1^(-1)`: element-wise reciprocal of obj1, i.e. `1/obj1`.
- `obj1[-1]`: drops the first element of obj1, returning the remaining 9 values.
- `obj2[4:2]`: returns obj2's 4th, 3rd, 2nd elements in that order, i.e. `7 5 3`.
- `obj1 + obj2`: obj2 (length 5) is recycled to length 10 to match obj1, then added element-wise (recycles cleanly since 10 is a multiple of 5, so no warning).
- `obj1 * obj3`: obj3 (logical, length 5) is coerced to numeric (TRUE=1, FALSE=0) and recycled to length 10, then multiplied element-wise.
- `obj2 + obj3`: obj3 coerced to numeric (0/0/1/1/1) and added element-wise to obj2.
- `obj1 + obj4`: obj4 is a list, not numeric, so this throws an error: `non-numeric argument to binary operator`.
- `sum(obj2)`: 2+3+5+7+11 = 28.
- `sum(obj3)`: logical TRUE/FALSE summed as 1/0 = 3.

## Question 2: Vectors

A user has typed the following commands into the console.

```{r}
a <- c(-1, 0)
b <- c(TRUE, FALSE)
c <- c("happy", "sad")
```

What do each of the following commands return? Pay attention to the type of the vector after concatenating two or more vectors.

```{r}
a+b
a*b
d <- c(a,b); d
e <- c(a,c); e
g <- c(b,c); g
w <- c(a,b,c); w
```

Notes:

- `a+b`: `b` is coerced to numeric (TRUE=1, FALSE=0), so `a+b = c(-1+1, 0+0) = c(0, 0)`.
- `a*b`: `a*b = c(-1*1, 0*0) = c(-1, 0)`.
- `d <- c(a,b)`: numeric + logical → coerced to numeric: `c(-1, 0, 1, 0)`.
- `e <- c(a,c)`: numeric + character → coerced to character: `c("-1", "0", "happy", "sad")`.
- `g <- c(b,c)`: logical + character → coerced to character: `c("TRUE", "FALSE", "happy", "sad")`.
- `w <- c(a,b,c)`: numeric + logical + character → all coerced to character (the most general type): `c("-1", "0", "TRUE", "FALSE", "happy", "sad")`.

## Question 3: Errors

For each of the following assignment statements, describe the error (or note why it does not generate an error).

```{r, eval=FALSE}
result1 <- floor 10.3
result2 <-- "Hello to you!"
result3 == "Hello to you!"
4result <- "Hello to you"
result5 <- "Hello to you
result6 <- date()
```

- `result1 <- floor 10.3`: **Syntax error.** `floor` is a function and requires parentheses around its argument (`floor(10.3)`). As written, R sees `floor` followed unexpectedly by the number `10.3` and reports "unexpected numeric constant."
- `result2 <-- "Hello to you!"`: **Runs, but errors at evaluation.** R parses this as `result2 <- -"Hello to you!"` (assignment followed by unary minus). You cannot apply unary minus to a character string, so R throws: "invalid argument to unary operator."
- `result3 == "Hello to you!"`: **Error, not an assignment.** `==` is the equality operator, not assignment, so this just tries to compare the (nonexistent) variable `result3` to a string. Since `result3` was never created, R reports "object 'result3' not found."
- `4result <- "Hello to you"`: **Syntax error.** Variable names in R cannot start with a digit, so `4result` is not a valid identifier ("unexpected symbol").
- `result5 <- "Hello to you`: **Syntax error.** The string is missing its closing quotation mark, so R reports an unterminated string constant / unexpected end of input.
- `result6 <- date()`: **No error.** `date()` is a valid built-in function that returns the current date and time as a character string (e.g. `"Wed Jul 01 2026 ..."`), which is assigned to `result6` successfully.

## Question 4: Matrices

#### 1) Create two matrices as follows by writing code. Call them `M1` and `M2`, respectively.

```{r}
a <- 1:9
M1 <- matrix(a,3,3)
M1
M2 <- matrix(a,3,3,byrow=TRUE)
M2
```

#### 2) Extract the value at the intersection of the 2nd row and 2nd column of M1 using R

```{r}
M1[2,2]
```

#### 3) Extract the values in the 3rd row in M1 using R

```{r}
M1[3,]
```

#### 4) Extract the values in the 3rd column in M2 using R

```{r}
M2[,3]
```

## Question 5: Calculations

#### 1) $log_{10}(1)+ln(1)$

```{r}
log10(1) + log(1)
```

#### 2) $sin(\pi/3)+cos(\pi/6)$

```{r}
sin(pi/3) + cos(pi/6)
```

#### 3) Round down the result of 2) to the nearest integer

```{r}
floor(sin(pi/3) + cos(pi/6))
```

#### 4) Round the result of 2) to four decimal places

```{r}
round(sin(pi/3) + cos(pi/6), 4)
```

#### 5) Calculate the remainder of $2024/3$ using R

```{r}
2024 %% 3
```

## Question 6: Data Frames

#### 1) Create a data frame as follows:

```{r}
Name <- c("Jacob","James","Jeremy","Jimmy","John")
Year <- c("1","2","1","4","3")
GPA <- c(3.41, 3.53,3.32,3.67,4.00)
Height <- c(1.78, 1.85, 1.72, 1.82, 1.68)
Students <- data.frame(Name,Year,GPA, Height)
Students
```

#### 2) Use the column name to extract the names of all students.

```{r}
Students$Name
```

#### 3) Select the information about James.

```{r}
Students[Students$Name == "James", ]
```

#### 4) List name(s) of students whose GPA is higher than 3.6.

```{r}
Students$Name[Students$GPA > 3.6]
```

#### 5) List the GPA of all freshmen and calculate the mean GPA.

```{r}
freshman_GPA <- Students$GPA[Students$Year == "1"]
freshman_GPA
mean(freshman_GPA)
```
