---
title: "R Homework"
output: html_document
---

## Question 1: Variables and Arithmetic (8 points)

Create the variables `x` and `y`:

```{r}
x <- 7
y <- 3
```

#### a. Compute $x^2 + y^2$

```{r}
x^2 + y^2
```

#### b. Compute the average of x and y

```{r}
(x + y) / 2
```

#### c. Determine whether x is greater than y

```{r}
x > y
```

## Question 2: Vectors (10 points)

```{r}
scores <- c(88, 91, 76, 95, 84, 90, 79)
scores
```

#### a. Find the mean.

```{r}
mean(scores)
```

#### b. Find the maximum value.

```{r}
max(scores)
```

#### c. Extract the first three elements.

```{r}
scores[1:3]
```

#### d. Replace the fourth score with 98.

```{r}
scores[4] <- 98
scores
```

#### e. Find all scores greater than 85.

```{r}
scores[scores > 85]
```

## Question 3: Data Frames (12 points)

```{r}
students <- data.frame(
  Name = c("Amy", "Ben", "Chris", "Dana"),
  Exam = c(88, 91, 76, 95),
  Homework = c(92, 85, 80, 98)
)
students
```

#### a. Display the first two rows.

```{r}
head(students, 2)
```

#### b. Compute the average exam score.

```{r}
mean(students$Exam)
```

#### c. Add a new column called "Total" equal to the average of Exam and Homework.

```{r}
students$Total <- (students$Exam + students$Homework) / 2
students
```

#### d. Return only the names of students with an Exam score above 90.

```{r}
students$Name[students$Exam > 90]
```

#### e. Return only the names of students with both Exam and Homework scores above 90.

```{r}
students$Name[students$Exam > 90 & students$Homework > 90]
```
